---
title: "Full Pedigree Analysis"
output: html_document
date: '2023-05-10'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Clean session
rm(list = ls())
install.packages("ggforce")
library(tidyverse)
library(ggplot2)
library(ggforce)
library(reshape2)
library(ggpubr)
```


##Plotting just the DNA tests and hereditary status
Insert the full pedigree with and without hereditary status
```{r}
PBGV_POAG <- read.csv("/Users/roscraddock/Documents/University of Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/FullCleanedData.csv")
```

#Pedigree visualisation - Full pedigree
```{r}
library(visPedigree)

FullPedDiag <- PBGV_POAG %>%
  select(IRN,IRN_sire,IRN_dam,Sex) %>%
  transform(Sex=ifelse(Sex =="Dog", "male", "female"))
FullPedDiag <- tidyped(FullPedDiag)
visped(FullPedDiag, compact = TRUE, cex = 0.83, file = "FullPedDiag.pdf")

```

Prepare the data for graphs
```{r}
rawData_DOB <- PBGV_POAG[!PBGV_POAG$DNA_result == "",]

# Change DOB format to date
rawData_DOB$DOB <- as.Date(rawData_DOB$DOB)

# Add year and month columns
rawData_DOB$year <- format(rawData_DOB$DOB, "%Y")
rawData_DOB$month <- format(rawData_DOB$DOB, "%m")
rawData_DOB$year_month <- format(rawData_DOB$DOB, "%Y-%m")
str(rawData_DOB)

#By year - using only DNA test and hereditary status
test <- rawData_DOB %>%
  select(year, DNA_result)

i <- 1995
df_DNA_DOB <- data.frame(matrix(nrow=0,ncol=4))
for (i in i: 2022){
  x <- test[test$year == i,]
  x <- x[!is.na(x$year),]
  t <- as.numeric(count(x))
  Clr <- as.numeric(count(x[x$DNA_result == "Clear" | x$DNA_result == "Hereditary Clear",]))
  Clr <- Clr/t
  Cr <- as.numeric(count(x[x$DNA_result == "Carrier" | x$DNA_result == "Hereditary Carrier",]))
  Cr <- Cr/t
  Af <- as.numeric(count(x[x$DNA_result == "Affected" | x$DNA_result == "Hereditary Affected",]))
  Af <- Af/t
  
  tmp <- cbind(i, Clr, Cr, Af)
  df_DNA_DOB <- rbind(df_DNA_DOB, tmp)
}

df_DNA_DOB <- na.omit(df_DNA_DOB)
df_DNA_DOB <- df_DNA_DOB %>%
  rename(Year = "i", AA = "Clr", AB = "Cr", BB="Af")

df_DNA_DOB <- melt(df_DNA_DOB, id.vars="Year", variable.name = "Genotype" )
```


Lets plot a graph with the data
```{r}
df_DNA_DOB %>% ggplot(aes(Year, value, group=Genotype, col=Genotype))+
  geom_line(size=1) +
  xlab("Year of Birth") +
  ylab("Mean Frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  scale_colour_manual(values = c("#009E73", "#0072B2", "#D55E00"), name = "Genotype", labels = c("AA", "AB", "BB"))

# ^^ not particularly informative due to the relatively low records per year. Especially earlier on.
```


Instead - lets plot the mutation allele frequency following Lewis and Mellersh
```{r}
test <- rawData_DOB %>%
  select(year, DNA_result)

i <- 1995
df_DNA_DOB <- data.frame(matrix(nrow=0,ncol=2))
df_T <- data.frame(matrix(nrow=0, ncol=2))
for (i in i: 2022){
  x <- test[test$year == i,]
  x <- x[!is.na(x$year),]
  t <- as.numeric(count(x))
  Clr <- as.numeric(count(x[x$DNA_result == "Clear" | x$DNA_result == "Hereditary Clear",]))
  Cr <- as.numeric(count(x[x$DNA_result == "Carrier" | x$DNA_result == "Hereditary Carrier",]))
  Af <- as.numeric(count(x[x$DNA_result == "Affected" | x$DNA_result == "Hereditary Affected",]))
  mf <- ((2*Af)+Cr)/t
  
  tmp <- cbind(i, mf)
  df_DNA_DOB <- rbind(df_DNA_DOB, tmp)
  tmp <- cbind(i, t)
  df_T <- rbind(df_T, tmp)
}

df_DNA_DOB <- na.omit(df_DNA_DOB)
df_DNA_DOB <- df_DNA_DOB %>%
  rename(Year = "i", Mutation_Frequency = "mf")
```

Plot the graph
```{r}
df_DNA_DOB %>% ggplot(aes(Year, Mutation_Frequency)) +
  geom_line(size=1, col = "#9834FF") +
  xlab("Year of Birth") +
  ylab("Mutation Frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))

#Restrict to 2000 in line with Lewis and Mellersh
test <- df_DNA_DOB[!df_DNA_DOB < 2000,]
test %>% ggplot(aes(Year, Mutation_Frequency)) +
  geom_line(size=1, col = "#9834FF") +
  xlab("Year of Birth") +
  ylab("Mutation Frequency") +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5))+
  geom_vline(xintercept = 2015, linetype = "dashed", colour = "#D35FB7")
```

Plotting the number of DNA tests and information per year
```{r}
ggplot(df_T, aes(i,t))+ 
  geom_point() +
  geom_line() +
  xlab("Year of Birth") +
  ylab("Number of Individuals with DNA Information")
```

Plotting the number of DNA tests per year since first published in 2015
```{r}

#Remove hereditary status
PBGV_POAG_DNA <- PBGV_POAG %>%
  transform(DNA_result=ifelse(DNA_result=="Hereditary Clear", NA, DNA_result))

PBGV_POAG_DNA <- PBGV_POAG_DNA %>%
  transform(DNA_result=ifelse(DNA_result=="Hereditary Carrier", NA, DNA_result))

PBGV_POAG_DNA <- PBGV_POAG_DNA %>%
  transform(DNA_result=ifelse(DNA_result=="Hereditary Affected", NA, DNA_result))

PBGV_POAG_DNA$DOB <- as.Date(PBGV_POAG_DNA$DOB)
PBGV_POAG_DNA$YearofBirth <- format(PBGV_POAG_DNA$DOB, "%Y")
DOB_PBGV <- PBGV_POAG_DNA[!is.na(PBGV_POAG_DNA$YearofBirth),]

DNA_Tests <- PBGV_POAG_DNA %>%
  select(DNA_result, DNA_TestDate)

DNA_Tests <- DNA_Tests[!DNA_Tests$DNA_result == "",]
DNA_Tests <- DNA_Tests[!is.na(DNA_Tests$DNA_result),]

# Change DOB format to date
DNA_Tests$DNA_TestDate <- as.Date(DNA_Tests$DNA_TestDate)

# Add year and month columns
DNA_Tests$year <- format(DNA_Tests$DNA_TestDate, "%Y")
DNA_Tests$year <- as.numeric(DNA_Tests$year)

Testperyear <- data.frame(matrix(nrow=0, ncol=2))
NumberBornperYear <- data.frame(matrix(nrow = 0, ncol = 2))
i <-1972
for(i in 1972:2022){
    x <- DNA_Tests %>%
    filter(year == i)
    z <- DOB_PBGV %>%
      filter(YearofBirth == i)
    no <- as.numeric(count(z))
    b <- cbind.data.frame(i, no)
    n <- as.numeric(count(x))
    y <- cbind.data.frame(i,n)
    Testperyear <- rbind(Testperyear, y)
    NumberBornperYear <- rbind(NumberBornperYear, b)
}

ggplot(Testperyear, aes(i,n))+ geom_point() +geom_line()+
  ylab("No. of DNA Tests")+xlab("Year")+xlim(c(2015,2022))

ggplot(NumberBornperYear, aes(i, no)) + geom_point() + geom_line()+
  ylab("Number of Dogs Born")+xlab("Year")+xlim(c(1972,2022))
```

## Full Pedigree from AlphaPeel
Insert the file with the genotype probabilities
```{r}
setwd(dir="/Users/roscraddock/Documents/Agriculture Uni/Year 3/NERC Placement/AlphaPeel Practice/AlphaPeel-master/AlphaPeel/outputs/Full Ped")

#Insert the haps (geno probabilities) file
GenoProbs_AlphaPeel <- read.csv("SingleLocus.haps", header = FALSE, sep = " ", col.names = c("IID", "Genotype_Probabilities"))

```

Adding the Genotype States
```{r}
#Add the genotype states/values
x <- c("AA", "AB", "BA", "BB")
x <- data.frame(x)
x <- x %>%
  rename(Geno_State = "x")
test <- x
ID <- ((as.numeric(count(GenoProbs_AlphaPeel)))/4)-1

for(i in 1:ID) {
test <- rbind(test, x)
}

GenoProbs_AlphaPeel <- cbind(GenoProbs_AlphaPeel, test)
```

Converting from tall to wide table format
```{r}
GenoProbs_AlphaPeel <- GenoProbs_AlphaPeel %>%
    pivot_wider(names_from = Geno_State,
                values_from = Genotype_Probabilities)

#Remove "mother of.." and "father of..."
GenoProbs_AlphaPeel$IID <- as.numeric(GenoProbs_AlphaPeel$IID)
GenoProbs_AlphaPeel <- GenoProbs_AlphaPeel %>%
  filter(!is.na(IID))

#Add in the DOB
Alpha_Peel_Recode <- read.csv("/Users/roscraddock/Documents/University\ of\ Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/fullPBGV_Alpha_Peel_Recode.csv")
Alpha_Peel_Recode <- Alpha_Peel_Recode %>%
  select(!X)

DOB <-Alpha_Peel_Recode %>%
  select(IID, DOB)

GenoProbs_AlphaPeel <- merge(GenoProbs_AlphaPeel, DOB, by="IID")
GenoProbs_AlphaPeel$IID <- as.numeric(GenoProbs_AlphaPeel$IID)

#Since the order of alleles is not relevant in this example, add columns Aa and aA together
GenoProbs_AlphaPeel$Het <- GenoProbs_AlphaPeel$BA + GenoProbs_AlphaPeel$AB
GenoProbsPhase_AlphaPeel <- GenoProbs_AlphaPeel
GenoProbs_AlphaPeel <- GenoProbs_AlphaPeel %>%
  select(IID, AA, Het, BB, DOB) %>%
  rename(AB = "Het")

```

##Plot the phased genotype probabilities averaged per birth year
```{r}
PhasedPlot <-GenoProbsPhase_AlphaPeel %>% 
  filter(!is.na(DOB))
#Removed 814 individuals

# Change DOB format to date
PhasedPlot$DOB <- as.Date(PhasedPlot$DOB)

# Add year and month columns
PhasedPlot$year <- format(PhasedPlot$DOB, "%Y")

#By year
DOB_Plot_Year <- PhasedPlot %>%
  group_by(year) %>%
  summarise(mean(AA)) 
DOB_Plot_Year2 <- PhasedPlot %>%
  group_by(year) %>%
  summarise(mean(AB))
DOB_Plot_Year3 <- PhasedPlot %>%
  group_by(year) %>%
  summarise(mean(BA))
DOB_Plot_Year4 <- PhasedPlot %>%
  group_by(year) %>%
  summarise(mean(BB))

PhasedPlot <- cbind(DOB_Plot_Year, DOB_Plot_Year2, DOB_Plot_Year3, DOB_Plot_Year4)

#Reduce to only one column for year
PhasedPlot <- PhasedPlot[,c(1,2,4,6, 8),]

PhasedPlot <- melt(PhasedPlot, id.vars="year", variable.name = "State" )

my_labels <- c("1976","", "1978", "", "1980", "", "1982", "", "1984", "", "1986", "", "1988", "","1990", "", "1992", "", "1994", "", "1996", "", "1998", "", "2000", "", "2002", "", "2004","", "2006", "", "2008", "", "2010", "", "2012", "", "2014","", "2016","", "2018","", "2020", "", "2022")

# Plotting the Mean frequency of states per year (1972 to January 2023)
PhasedPlot %>% ggplot(aes(year, value, group=State, col=State))+
  geom_line(size=1) +
  xlab("Year of Birth") +
  ylab("Population Frequency") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5)) +
  scale_x_discrete(breaks = seq(1976, 2022), labels = my_labels) +
  scale_colour_manual(values = c("#009E73", "#0072B2", "#56B4E9", "#D55E00"), name = "Genotype", labels = c("AA", "AB", "BA", "BB"))+
  geom_vline(xintercept = "1995", linetype = "dotted") +
  geom_vline(xintercept = "2015", linetype = "dashed", colour = "#D35FB7")

```

## Plot the genotype probabilities averaged per birth year.
```{r}
#Remove individuals with no recorded date of birth
DOB_Plot <- GenoProbs_AlphaPeel %>% 
  filter(!is.na(DOB))
#Removed 814 individuals

# Change DOB format to date
DOB_Plot$DOB <- as.Date(DOB_Plot$DOB)

# Add year and month columns
DOB_Plot$year <- format(DOB_Plot$DOB, "%Y")
DOB_Plot$month <- format(DOB_Plot$DOB, "%m")
DOB_Plot$year_month <- format(DOB_Plot$DOB, "%Y-%m")
str(DOB_Plot)

#By year
DOB_Plot_Year <- DOB_Plot %>%
  group_by(year) %>%
  summarise(mean(AA)) 
DOB_Plot_Year2 <- DOB_Plot %>%
  group_by(year) %>%
  summarise(mean(AB))
DOB_Plot_Year3 <- DOB_Plot %>%
  group_by(year) %>%
  summarise(mean(BB))

DOB_Plot_Year <- cbind(DOB_Plot_Year, DOB_Plot_Year2, DOB_Plot_Year3)

#Reduce to only one column for year
DOB_Plot_Year <- DOB_Plot_Year[,c(1,2,4,6),]

#Move back into the tall format
DOB_Plot_Year <- melt(DOB_Plot_Year, id.vars="year", variable.name = "State" )

my_labels <- c("1976","", "1978", "", "1980", "", "1982", "", "1984", "", "1986", "", "1988", "","1990", "", "1992", "", "1994", "", "1996", "", "1998", "", "2000", "", "2002", "", "2004","", "2006", "", "2008", "", "2010", "", "2012", "", "2014","", "2016","", "2018","", "2020", "", "2022")

# Plotting the Mean frequency of states per year (1972 to January 2023)
DOB_Plot_Year %>% ggplot(aes(year, value, group=State, col=State))+
  geom_line(size=1) +
  xlab("Year of Birth") +
  ylab("Population Frequency") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5)) +
  scale_x_discrete(breaks = seq(1976, 2022), labels = my_labels) +
  scale_colour_manual(values = c("#009E73", "#0072B2", "#D55E00"), name = "Genotype", labels = c("AA", "AB", "BB"))+
  geom_vline(xintercept = "1995", linetype = "dotted") +
  geom_vline(xintercept = "2015", linetype = "dashed", colour = "#D35FB7")

#breaks=seq(1976, 2022, 2)
#Plot by Mean frequency of states per year from 1995 (earliest DOB with DNA test)
test <- DOB_Plot_Year[!DOB_Plot_Year$year < 1995,]
test$State <- as.character(test$State)
test %>% ggplot(aes(year, value, group=State, col=State))+
  geom_line(size=1) +
  xlab("Year of Birth") +
  ylab("Population Frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  scale_colour_manual(values = c("#009E73", "#0072B2", "#D55E00"), name = "Genotype", labels = c("AA", "AB", "BB")) +
  geom_vline(xintercept = "2015", linetype = "dashed", colour = "#D35FB7")

#Save a graph for dates 2009 to June 2014
test <- DOB_Plot[!DOB_Plot$DOB < "2008-12-31",]
test <- test[!test$DOB > "2014-06-30",]

#By year
tmp <- test %>%
  group_by(year) %>%
  summarise(mean(AA))
tmp2 <- test %>%
  group_by(year) %>%
  summarise(mean(AB))
tmp3 <- test %>%
  group_by(year) %>%
  summarise(mean(BB))

Full_Sub <- cbind(tmp, tmp2, tmp3)

#Reduce to only one column for year
Full_Sub <- Full_Sub[,c(1,2,4,6),]

#Move back into the tall format
Full_Sub <- melt(Full_Sub, id.vars="year", variable.name = "State" )

# Plotting the Mean frequency of states per year (1972 to January 2023)
FullSubPlot <- Full_Sub %>% ggplot(aes(year, value, group=State, col=State))+
  geom_line(size=1) +
  xlab("Year of Birth") +
  ylab("Population Frequency") +
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5)) +
  scale_colour_manual(values = c("#009E73", "#0072B2", "#D55E00"), name = "Genotype", labels = c("AA", "AB", "BB")) + ylim(0,1)


```

#Add a line for the mean minor allele frequency
```{r}
#First Find the minor allele frequency per year using all the individuals
DOB_Plot_MA <- DOB_Plot_Year %>%
  pivot_wider(names_from = "State", values_from = "value")
  
DOB_Plot_MA$MAF <- DOB_Plot_MA$`mean(BB)`+ DOB_Plot_MA$`mean(AB)`* 0.5

DOB_Plot_MA <- melt(DOB_Plot_MA, id.vars="year", variable.name = "State" )

DOB_Plot_MA$State <- factor(DOB_Plot_MA$State, levels = c("mean(AA)", "mean(AB)", "mean(BB)", "MAF"))

# Plotting the Mean frequency of states per year (1972 to January 2023
TP <- DOB_Plot_MA %>% ggplot(aes(year, value, group=State, col=State))+
  geom_line(size=1) +
  xlab("Year of Birth") +
  ylab("Mean Frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  scale_colour_manual(values = c("#009E73", "#0072B2", "#D55E00", "#7C788A"), name = "Genotype", labels = c("AA", "AB", "BB", "Mutation Allele (B)"))+
  geom_vline(xintercept = "1995", linetype = "dotted") +
  geom_vline(xintercept = "2015", linetype = "dashed", colour = "#D35FB7")

ggarrange(TP, nrow = 1, ncol = 1, legend = "none")

#Plot by Mean frequency of states per year from 2010 (When DNA information per year exceeds 30)
test <- DOB_Plot_MA[!DOB_Plot_MA$year < 2010,]
test$State <- as.character(test$State)
test$State <- factor(test$State, levels = c("mean(AA)", "mean(AB)", "mean(BB)", "MAF"))
test %>% ggplot(aes(year, value, group=State, col=State))+
  geom_line(size=1) +
  xlab("Year of Birth") +
  ylab("Mean Frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  scale_colour_manual(values = c("#009E73", "#0072B2", "#D55E00", "#7C788A"), name = "Genotype", labels = c("AA", "AB", "BB", "Mutation Allele Frequency")) +
  geom_vline(xintercept = "2015", linetype = "dashed", colour = "#D35FB7")

#Plot the same time line with the estimated mutation frequencies from DNA test and hereditary status
test <- df_DNA_DOB[!df_DNA_DOB < 2010,]
test %>% ggplot(aes(Year, Mutation_Frequency)) +
  geom_line(size=1, col = "#9834FF") +
  xlab("Year of Birth") +
  ylab("Mutation Frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))


```

#Add in the prediction using DNA test and hereditary status only
```{r}
DOB_Plot_MA <- DOB_Plot_MA %>%
  pivot_wider(names_from = "State", values_from = "value")

DOB_Plot_MA_Wide <- merge( DOB_Plot_MA, df_DNA_DOB, by.x = "year", by.y = "Year", all.x = TRUE)


```

#Add minor allele frequency from DNA test and hereditary status to the above graph
```{r}
test <- melt(DOB_Plot_MA_Wide, id.vars = "year", variable.name = "State")
test <- test[!test$year < 2000,]
test$State <- factor(test$State, levels = c("mean(AA)", "mean(AB)", "mean(BB)", "MAF", "Mutation_Frequency"))

test %>% ggplot(aes(year, value, group=State, col=State))+
  geom_line(size=1) +
  xlab("Year of Birth") +
  ylab("Population Frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  scale_colour_manual(values = c("#009E73", "#0072B2", "#D55E00", "#7C788A", "#9834FF"), name = "Genotype", labels = c("AA", "AB", "BB", "Imputed Mutation Allele Frequency", "DNA Information Mutation Allele Frequency")) +
  geom_vline(xintercept = "2015", linetype = "dashed", colour = "#D35FB7")
```

#Compare with the minor allele frequency from the DNA test and hereditary data.
```{r}

test <- DOB_Plot_MA_Wide %>%
  select(year, MAF, Mutation_Frequency)

test <- melt(test, id.vars = "year", variable.name = "State")

test <- test[test$year > 1999,]
test$State <- factor(test$State, levels = c("Mutation_Frequency", "MAF"))


test %>% ggplot(aes(year, value, group=State, col=State))+
  geom_line(size=1) +
  xlab("Year of Birth") +
  ylab("Estimated Mutation Frequency") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 0.5)) +
  scale_colour_manual(values = c("#9834FF","#7C788A"), name = "", labels = c("DNA Information", "Full Pedigree Imputation")) +
  geom_vline(xintercept = "2015", linetype = "dashed", colour = "#D35FB7")

```

#Graph with the imputed minor and major allele frequencies
```{r}
DOB_Plot_MA$MjAF <- 1 - DOB_Plot_MA$MAF

test <- DOB_Plot_MA %>%
  select(year, MAF, MjAF)

test <- melt(test, id.vars = "year", variable.name = "State")
test$State <- factor(test$State, levels = c("MAF", "MjAF"))

test %>% ggplot(aes(year, value, group=State, col=State))+
  geom_line(size=1) +
  xlab("Year of Birth") +
  ylab("Estimated Allele Frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  scale_colour_manual(values = c("#7C788A", "#44AA99"), name = "", labels = c("Disease-causing Allele (B)", "Normal Allele (A)")) +
  geom_vline(xintercept = "2015", linetype = "dashed", colour = "#D35FB7")

```

#Validation of full pedigree imputation in AlphaPeel - comparing hereditary status with genotype dosages by pearson coefficient
```{r}

#Insert full pedigree with DNA test and hereditary status in the same order as AlphaPeel
Val_Alpha_Peel <- read_csv("/Users/roscraddock/Documents/PBGV_Full_Val_Alpha.csv")
#Insert the genotype dosages
FullPed_GenDos <- read.csv("/Users/roscraddock/Documents/Agriculture Uni/Year 3/NERC Placement/AlphaPeel Practice/AlphaPeel-master/AlphaPeel/outputs/Full Ped/SingleLocus.dosages", header = FALSE, sep = " ", col.names = c("IID", "Genotype_Dosage"))
#Merge Val_Alpha with dosages
Val_Alpha_Peel <- merge(FullPed_GenDos, Val_Alpha_Peel, by="IID")

#Select only those with hereditary status
tmp <- Val_Alpha_Peel %>%
  filter(Genotype =="Hereditary Clear") %>%
  transform(Genotype = 0)
tmp1 <- Val_Alpha_Peel %>%
  filter(Genotype == "Hereditary Carrier")%>%
  transform(Genotype = 1)
tmp2 <- Val_Alpha_Peel %>%
  filter(Genotype == "Hereditary Affected") %>%
  transform(Genotype = 2)

Val_Alpha_Peel <- rbind(tmp, tmp1, tmp2)

Val_Alpha_Peel <- Val_Alpha_Peel %>%
  select(IID, Genotype_Dosage, Genotype, DOB)


#For all hereditary status (374 individuals)
Val_Alpha_Peel$Genotype <- as.numeric(Val_Alpha_Peel$Genotype)
cor <- cor.test(Val_Alpha_Peel$Genotype_Dosage, Val_Alpha_Peel$Genotype, method = "pearson")

Val_Alpha_Peel$Genotype <- as.character(Val_Alpha_Peel$Genotype)

Val_Alpha_Peel %>% ggplot() + aes(Genotype, Genotype_Dosage)+ geom_sina() +
  ylab("Imputed Genotype Dosages") + xlab("Genotype Dosages from Hereditary Status")


#For those in 2022
Val_Alpha_Peel$DOB <- as.Date(Val_Alpha_Peel$DOB, "%Y-%m-%d")
Val_Alpha_Peel$year <- format(Val_Alpha_Peel$DOB, "%Y")
Val_Alpha_Peel_2022 <- Val_Alpha_Peel[Val_Alpha_Peel$year == 2022,]
Val_Alpha_Peel_2022$Genotype <- as.numeric(Val_Alpha_Peel_2022$Genotype)
Val_Alpha_Peel_2022 <- Val_Alpha_Peel_2022[!is.na(Val_Alpha_Peel_2022$DOB),]
cor <- cor(Val_Alpha_Peel_2022$Genotype, Val_Alpha_Peel_2022$Genotype_Dosage, method = "pearson")

Val_Alpha_Peel_2022$Genotype <- as.character(Val_Alpha_Peel_2022$Genotype)

Val_Alpha_Peel_2022 %>% ggplot() + aes(Genotype, Genotype_Dosage)+ geom_sina() +
  ylab("Imputed Genotype Dosages") + xlab("Genotype Dosages from Hereditary Status")

```

