---
title: "Sub-Pedigree LOOCV"
output: html_document
date: '2023-05-10'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Clean session
rm(list = ls())
#Load in all packages here.
# Install
if (FALSE) {
  setRepositories() # choose CRAN and BioC
  install.packages(pkg = c("gRain", "devtools"), dep = TRUE)
  # PR that fixed the repeatPattern() data issue
  # https://github.com/hojsgaard/gRain/pull/9
  devtools::install_github(repo = "https://github.com/gregorgorjanc/gRain")
}
# Load package
library(package = "gRain")

# User defined functions

source(file = "AlleleFetch.R")

library(tidyverse)
library(ggplot2)
```

## Start with Genotype and Allelic Concordance in AlphaPeel
First need to import all masked genotype runs (127) from AlphaPeel.

```{r}

#Import AlphaGeno file
Alpha_Peel_Geno <- read.csv("/Users/roscraddock/Documents/University of Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/AlphaPeel Inputs/PBVG_Geno_2009-2014.txt", sep = "", col.names = c("IID", "Genotype"), header = FALSE)

#Set the working directory for this section of r
setwd(dir="/Users/roscraddock/Documents/Agriculture Uni/Year 3/NERC Placement/AlphaPeel Practice/AlphaPeel-master/AlphaPeel/outputs")

#Separate only those with a genotype recorded
Geno <- Alpha_Peel_Geno[!Alpha_Peel_Geno$Genotype==9,]

#For loop to insert imputed mask genotype into a dataframe
i <- 5
Imputed_Geno <- data.frame (matrix(nrow = 0, ncol = 2))
df.GenoConcordance <- data.frame(matrix(nrow = 0, ncol = 3))
nID <- as.numeric(count(Alpha_Peel_Geno))
for (i in i:nID){
  test <- i %in% Geno$IID
  if(any(test)){
    Filename <- paste(i, "SingleLocus2009-2014.haps", sep="")
    tmp <- read.csv(Filename, sep = " ", col.names = c("IID", "Genotype_Probabilities"), header = FALSE)
    tmp <- tmp %>%
      filter(IID == i)
    tmp$Genotype_Probabilities <- as.numeric(tmp$Genotype_Probabilities)
    y <- tmp[2,2,]
    z <- tmp[3,2,]
    w <- y+z
  
    tmp$Genotype_Probabilities[2] <- w
    tmp <- tmp[-3,]
  
    Imputed_Geno <- rbind(Imputed_Geno, tmp)
   
   #Genotype Concordance Set Up
   ID <- as.numeric(Geno %>%
     select(IID) %>%
     filter(IID == i))
  Genotype <- Geno %>%
    filter(IID == i) %>%
    select(Genotype)
    if (Genotype == 0) {
    GenProb <- structure(list(True= c(1, 0, 0)), class = "data.frame", row.names = c(NA, -3))
    } 
    if (Genotype == 1) {
    GenProb <- structure(list(True= c(0, 1, 0)), class = "data.frame", row.names = c(NA, -3))
    } 
    if (Genotype == 2) {
    GenProb <- structure(list(True= c(0, 0, 1)), class = "data.frame", row.names = c(NA, -3))
    } 
  
  GenoCon <- tmp$Genotype_Probabilities * GenProb
  GenoConcordance <- sum(GenoCon)
  
  #Allelic Concordance
  if (Genotype == 0){
    AM <- structure(list(True= c(1, 0.5, 0)), class = "data.frame", row.names = c(NA, -3))
  }
  if (Genotype == 1){
    AM <- structure(list(True= c(0.5, 1, 0.5)), class = "data.frame", row.names = c(NA, -3))
  }
  if (Genotype == 2){
    AM <- structure(list(True= c(0, 0.5, 1)), class = "data.frame", row.names = c(NA, -3))
  }

  AlleCon <- tmp$Genotype_Probabilities * AM
  AllelicConcordance <- sum(AlleCon)
  #Insert table to add to.
  tmp2 <- cbind(i, GenoConcordance, AllelicConcordance)
  df.GenoConcordance <- rbind(df.GenoConcordance, tmp2)
  
  }
}

#Add the genotype states/values
x <- c("AA", "AB", "BB")
x <- data.frame(x)
x <- x %>%
  rename(Geno_State = "x")
test <- x
ID <- ((as.numeric(count(Imputed_Geno)))/3)-1

for(i in 1:ID) {
test <- rbind(test, x)
}

Imputed_Geno <- cbind(Imputed_Geno, test)
```

#Pearson coefficicent between observed and imputed when masked in AlphaPeel
```{r}

#First widen imputed_geno
df_ImputedGeno <- Imputed_Geno %>%
  pivot_wider(names_from = Geno_State, values_from = Genotype_Probabilities)
#Add column with the geno dosage
df_ImputedGeno$ImpGenoDosage <- df_ImputedGeno$AA*0+df_ImputedGeno$AB*1+df_ImputedGeno$BB*2
#Merge with Geno
df_ImputedGeno <- merge(df_ImputedGeno, Geno, by = "IID")
#Pearson correlation
CorAP <- cor.test (df_ImputedGeno$ImpGenoDosage, df_ImputedGeno$Genotype, method = "pearson")
```

#Summary statistics
```{r}
TO DO
```


## LOOCV in AlleleFetch for the Three Models

## Importing and preparing the data
```{r}
#Insert the datasets
GenoScale <- read.csv("/Users/roscraddock/Documents/University of Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/AlleleFetch Inputs/PBGV_Geno_2009-2014.csv")
GenoScale <- GenoScale %>%
  select(!X)
PhenoScale <- read.csv("/Users/roscraddock/Documents/University of Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/AlleleFetch Inputs/PBGV_Pheno_2009-2014.csv")
PhenoScale <- PhenoScale %>%
  select(!X)
PedScale <- read.csv("/Users/roscraddock/Documents/University of Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/AlleleFetch Inputs/PBGV_Ped_2009-2014.csv")
PedScale <- PedScale %>%
  select(!X)
PBGV_POAG_Sub <- read.csv("/Users/roscraddock/Documents/University\ of\ Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/PBGV_Sub_2009-2014.csv")
PBGV_POAG_Sub <- PBGV_POAG_Sub %>%
  select(!X)
```


##Setting up the parameters for AlleleFetch
```{r}
#Setting up the genotype and phenotype states
testGenotypeState <- c("AA", "AB", "BB")
testPhenotypeState <- c("OK", "NOK")

# Prior allele frequency in (founder) population - #0.309474 taken from AlphaPeel
PrB <- 0.309474 
PrA <- 1-PrB


# Prior genotype frequencies in (founder) population
# according to the Hardy-Weinberg's expectation
testGenoPT <- c(PrA*PrA, 2*PrA*PrB, PrB*PrB)

# Genotype frequencies in individual given the genotype of parents according to Mendel's law
# and accounting for mutation (via e)
e <- 0.00000001
gM_AA_AA <- c(1-2*e+e^2, 2*e-2*e^2, e^2)
gM_AA_AB <- c(1/2-1/2*e, 1/2, 1/2*e)
gM_AA_BB <- c(e-e^2, 1-2*e+2*e^2, e-e^2)
gM_AB_AA <- c(1/2-1/2*e, 1/2, 1/2*e)
gM_AB_AB <- c(1/4, 1/2, 1/4)
gM_AB_BB <- c(1/2*e, 1/2, 1/2-1/2*e)
gM_BB_AA <- c(e-e^2, 1-2*e+2*e^2, e-e^2)
gM_BB_AB <- c(1/2*e, 1/2, 1/2-1/2*e)
gM_BB_BB <- c(e^2, 2*e-2*e^2, 1-2*e+e^2)

testMendelianCPT <- matrix(nrow=3, ncol=9)
testMendelianCPT[, 1] <- gM_AA_AA
testMendelianCPT[, 2] <- gM_AA_AB
testMendelianCPT[, 3] <- gM_AA_BB
testMendelianCPT[, 4] <- gM_AB_AA
testMendelianCPT[, 5] <- gM_AB_AB
testMendelianCPT[, 6] <- gM_AB_BB
testMendelianCPT[, 7] <- gM_BB_AA
testMendelianCPT[, 8] <- gM_BB_AB
testMendelianCPT[, 9] <- gM_BB_BB

# Genotype frequencies in individual given the genotype of one parent (Mendel's law with mutation)
# and base population (Hardy-Weinberg's law)
gM_AA_unk <- c(PrA-PrA*e, PrB-PrB*e+PrA*e, PrB*e)
gM_AB_unk <- c(1/2*PrA, 1/2*(PrA+PrB), 1/2*PrB)
gM_BB_unk <- c(PrA*e, PrA-PrA*e+PrB*e, PrB-PrB*e)

testMendelianCPTOnlyFather <- matrix(nrow=3, ncol= 3)
testMendelianCPTOnlyFather[,1] <- gM_AA_unk
testMendelianCPTOnlyFather[,2] <- gM_AB_unk
testMendelianCPTOnlyFather[,3] <- gM_BB_unk

testMendelianCPTOnlyMother <- testMendelianCPTOnlyFather

# Set of phenotype values
testPhenotypeState <- c("OK", "NOK")

# Phenotype frequences in individual given the genotype of the individual - phenotype penetrance
e <- 0.01
testPhenoPenetranceCPT <- matrix(nrow=2, ncol=3)
#                               OK   NOK Pheno
testPhenoPenetranceCPT[,1] <- c(1-e, 0+e) # AA True geno
testPhenoPenetranceCPT[,2] <- c(1-e, 0+e) # AB
testPhenoPenetranceCPT[,3] <- c(0+e, 1-e) # BB   

# Genotype frequences in individual given the genotype of the individual - genotype penetrance
e <- 0.1
testGenoPenetranceCPT <- matrix(nrow=3, ncol=3)
#                             AA       AB     BB Observed Geno
testGenoPenetranceCPT[,1] <-c(1-e-e^2, e,     e^2)     # AA True Geno
testGenoPenetranceCPT[,2] <-c(e,       1-2*e, e)       # AB
testGenoPenetranceCPT[,3] <-c(e^2,     e,     1-e-e^2) # BB
```

#Using LOOCV to collect genotype and allelic concordance for Genotype only
```{r}

#Start a table to add concordance to (with True genotypes)
col <- c("IID", "DNA Test", "Genotype Concordance", "Allelic Concordance", "Imputation Dosage", "AA", "AB", "BB")
df_geno_only <- data.frame(matrix(nrow=0,ncol=length(col)))
colnames(df_geno_only)=col

#Set up count for loop
i <- 1
nID <- count(GenoScale)
#"True Genotype" result as genotype probabilities
for(i in 1:nID$n) {
  ID <- GenoScale[i,1,]
  Genotype <- GenoScale[i,2,]
  if (Genotype == "AA") {
  GenProb <- structure(list(True= c(1, 0, 0)), class = "data.frame", row.names = c(NA, -3))
  } 
  if (Genotype == "AB") {
  GenProb <- structure(list(True= c(0, 1, 0)), class = "data.frame", row.names = c(NA, -3))
  } 
  if (Genotype == "BB") {
  GenProb <- structure(list(True= c(0, 0, 1)), class = "data.frame", row.names = c(NA, -3))
  } 
  
    
  #Mask the DNA result by removing before running
  GenoMask <- GenoScale[!GenoScale$IID==ID,]
  #Run AlleleFetch
  out <- AlleleFetch(pedigreeData = PedScale,
                   genotypeState = testGenotypeState,
                  genotypeData = GenoMask,
                   phenotypeState = testPhenotypeState,
                   genoPT = testGenoPT,
                   mendelianCPT = testMendelianCPT,
                   mendelianCPTOnlyFather = testMendelianCPTOnlyFather,
                   mendelianCPTOnlyMother = testMendelianCPTOnlyMother,
                   genoPenetranceCPT = testGenoPenetranceCPT,
                   phenoPenetranceCPT = testPhenoPenetranceCPT)
  #Save results into dataframe
  Result <- out$result
  
  tmp1 <- Result %>%
    filter(IID == ID) %>%
    filter(State == "AA") %>%
    select(State, Freq)
  tmp2 <- Result %>%
    filter(IID == ID) %>%
    filter(State == "AB") %>%
    select(State, Freq)
  tmp3 <- Result %>%
    filter(IID == ID) %>%
    filter(State == "BB") %>%
    select(State, Freq)
  PostGeno <- rbind(tmp1, tmp2, tmp3)
  Imp_Geno_Dosage <- 0*tmp1$Freq + 1*tmp2$Freq +2*tmp3$Freq
  
  #Genotype Concordance
  GenoCon <- PostGeno$Freq * GenProb
  GenoConcordance <- sum(GenoCon)

  
  #Allelic Concordance
  if (Genotype == "AA"){
    AM <- structure(list(True= c(1, 0.5, 0)), class = "data.frame", row.names = c(NA, -3))
  }
  if (Genotype == "AB"){
    AM <- structure(list(True= c(0.5, 1, 0.5)), class = "data.frame", row.names = c(NA, -3))
  }
  if (Genotype == "BB"){
    AM <- structure(list(True= c(0, 0.5, 1)), class = "data.frame", row.names = c(NA, -3))
  }

  AlleCon <- PostGeno$Freq * AM
  AllelicConcordance <- sum(AlleCon)
  #Insert table to add to.
  tmp <- cbind(ID, Genotype, GenoConcordance, AllelicConcordance, Imp_Geno_Dosage, tmp1$Freq, tmp2$Freq, tmp3$Freq)
  df_geno_only <- rbind(df_geno_only, tmp)
  
}
```

#Using LOOCV to collect genotype and allelic concordance for observed genotype with observed phenotypes
```{r}

 #Start a table to add concordance to (with True genotypes)
col <- c("IID", "DNA Test", "Genotype Concordance", "Allelic Concordance","AA", "AB", "BB")
df_pheno <- data.frame(matrix(nrow=0,ncol=length(col)))
colnames(df_pheno)=col

#Set up count for loop
nID <- count(GenoScale)
#"True Genotype" result as genotype probabilities
for(i in 1:nID$n) {
  ID <- GenoScale[i,1,]
  Genotype <- GenoScale[i,2,]
  if (Genotype == "AA") {
  GenProb <- structure(list(True= c(1, 0, 0)), class = "data.frame", row.names = c(NA, -3))
  } 
  if (Genotype == "AB") {
  GenProb <- structure(list(True= c(0, 1, 0)), class = "data.frame", row.names = c(NA, -3))
  } 
  if (Genotype == "BB") {
  GenProb <- structure(list(True= c(0, 0, 1)), class = "data.frame", row.names = c(NA, -3))
  } 
  
    
  #Mask the DNA result by removing before running
  GenoMask <- GenoScale[!GenoScale$IID==ID,]
  #Run AlleleFetch
  out <- AlleleFetch(pedigreeData = PedScale,
                   genotypeState = testGenotypeState,
                   genotypeData = GenoMask,
                   phenotypeState = testPhenotypeState,
                   phenotypeData = PhenoScale,
                   genoPT = testGenoPT,
                   mendelianCPT = testMendelianCPT,
                   mendelianCPTOnlyFather = testMendelianCPTOnlyFather,
                   mendelianCPTOnlyMother = testMendelianCPTOnlyMother,
                   genoPenetranceCPT = testGenoPenetranceCPT,
                   phenoPenetranceCPT = testPhenoPenetranceCPT)
  #Save results into dataframe
  Result <- out$result
  
  tmp1 <- Result %>%
    filter(IID == ID) %>%
    filter(State == "AA") %>%
    select(State, Freq)
  tmp2 <- Result %>%
    filter(IID == ID) %>%
    filter(State == "AB") %>%
    select(State, Freq)
  tmp3 <- Result %>%
    filter(IID == ID) %>%
    filter(State == "BB") %>%
    select(State, Freq)
  PostGeno <- rbind(tmp1, tmp2, tmp3)
  Imp_Geno_Dosage <- 0*tmp1$Freq + 1*tmp2$Freq +2*tmp3$Freq 
  
  #Genotype Concordance
  GenoCon <- PostGeno$Freq * GenProb
  GenoConcordance <- sum(GenoCon)

  
  #Allelic Concordance
  if (Genotype == "AA"){
    AM <- structure(list(True= c(1, 0.5, 0)), class = "data.frame", row.names = c(NA, -3))
  }
  if (Genotype == "AB"){
    AM <- structure(list(True= c(0.5, 1, 0.5)), class = "data.frame", row.names = c(NA, -3))
  }
  if (Genotype == "BB"){
    AM <- structure(list(True= c(0, 0.5, 1)), class = "data.frame", row.names = c(NA, -3))
  }

  AlleCon <- PostGeno$Freq * AM
  AllelicConcordance <- sum(AlleCon)
  #Insert table to add to.
  tmp <- cbind(ID, Genotype, GenoConcordance, AllelicConcordance, Imp_Geno_Dosage, tmp1$Freq, tmp2$Freq, tmp3$Freq)
  df_pheno <- rbind(df_pheno, tmp)
  
}

```

Save file to prevent loss/deletion
```{r}
write.csv(df_geno_only, "Geno_13-18_AFAP_LowMutationrate.csv")
write.csv(df_pheno, "Pheno_13-18_AFAP_LowMutationrate.csv")
```

## To test/observe any discordance
```{r}
test <- merge(PhenoScale, GenoScale, by="IID")
test <- merge(test, AgePhenoScale, by = "IID")
table(test$Genotype, test$Phenotype)
```

## Introduction of new penetrance function that accounts for age
```{r}

#Additional phenotype function for age at eye test under 4.
testPhenoPenetranceCPT_Before <- matrix(nrow=2, ncol=3)
 #                           OK<4 NOK<4 Pheno
 testPhenoPenetranceCPT_Before[,1] <- c(1-e, 0+e) # AA True geno
 testPhenoPenetranceCPT_Before[,2] <- c(1-e, 0+e) # AB
 testPhenoPenetranceCPT_Before[,3] <- c(1-e, 0+e) # BB
 
 #Additional dataset for the age at eye testing
PBGV_POAG_Sub <- read.csv("/Users/roscraddock/Documents/University\ of\ Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/PBGV_Sub_2009-2014.csv")

AgePhenoScale <- PBGV_POAG_Sub %>%
  select(IRN, DOB, EyeTest_Date)
AgePhenoScale <- AgePhenoScale %>%
  filter(!EyeTest_Date == "")
AgePhenoScale$DOB <- as.Date(AgePhenoScale$DOB)
AgePhenoScale$EyeTest_Date <- as.Date(AgePhenoScale$EyeTest_Date)

AgePhenoScale$Age <- AgePhenoScale$EyeTest_Date - AgePhenoScale$DOB
AgePhenoScale$Age <- as.numeric(AgePhenoScale$Age) %/% 365.25

AgePhenoScale <- AgePhenoScale %>%
  select(IRN, Age) %>%
  rename(IID = "IRN")
```


#Using LOOCV to collect genotype and allelic concordance for observed genotype with observed phenotypes with age penetrance function
```{r}


 #Start a table to add concordance to (with True genotypes)
col <- c("IID", "DNA Test", "Genotype Concordance", "Allelic Concordance","AA", "AB", "BB")
df_pheno_Pen <- data.frame(matrix(nrow=0,ncol=length(col)))
colnames(df_pheno_Pen)=col

#Set up count for loop
nID <- count(GenoScale)
#"True Genotype" result as genotype probabilities
for(i in 1:nID$n) {
  ID <- GenoScale[i,1,]
  Genotype <- GenoScale[i,2,]
  if (Genotype == "AA") {
  GenProb <- structure(list(True= c(1, 0, 0)), class = "data.frame", row.names = c(NA, -3))
  } 
  if (Genotype == "AB") {
  GenProb <- structure(list(True= c(0, 1, 0)), class = "data.frame", row.names = c(NA, -3))
  } 
  if (Genotype == "BB") {
  GenProb <- structure(list(True= c(0, 0, 1)), class = "data.frame", row.names = c(NA, -3))
  } 
  
    
  #Mask the DNA result by removing before running
  GenoMask <- GenoScale[!GenoScale$IID==ID,]
  #Run AlleleFetch
  out <- AlleleFetch(pedigreeData = PedScale,
                   genotypeState = testGenotypeState,
                   genotypeData = GenoMask,
                   phenotypeState = testPhenotypeState,
                   phenotypeData = PhenoScale,
                   genoPT = testGenoPT,
                   mendelianCPT = testMendelianCPT,
                   mendelianCPTOnlyFather = testMendelianCPTOnlyFather,
                   mendelianCPTOnlyMother = testMendelianCPTOnlyMother,
                   genoPenetranceCPT = testGenoPenetranceCPT,
                   phenoPenetranceCPT = testPhenoPenetranceCPT,
                   phenoPenetranceCPT_Before = testPhenoPenetranceCPT_Before,
                   agePhenotype = AgePhenoScale)
  #Save results into dataframe
  Result <- out$result
  
  tmp1 <- Result %>%
    filter(IID == ID) %>%
    filter(State == "AA") %>%
    select(State, Freq)
  tmp2 <- Result %>%
    filter(IID == ID) %>%
    filter(State == "AB") %>%
    select(State, Freq)
  tmp3 <- Result %>%
    filter(IID == ID) %>%
    filter(State == "BB") %>%
    select(State, Freq)
  PostGeno <- rbind(tmp1, tmp2, tmp3)
  Imp_Geno_Dosage <- 0*tmp1$Freq + 1*tmp2$Freq +2*tmp3$Freq
  
  #Genotype Concordance
  GenoCon <- PostGeno$Freq * GenProb
  GenoConcordance <- sum(GenoCon)

  
  #Allelic Concordance
  if (Genotype == "AA"){
    AM <- structure(list(True= c(1, 0.5, 0)), class = "data.frame", row.names = c(NA, -3))
  }
  if (Genotype == "AB"){
    AM <- structure(list(True= c(0.5, 1, 0.5)), class = "data.frame", row.names = c(NA, -3))
  }
  if (Genotype == "BB"){
    AM <- structure(list(True= c(0, 0.5, 1)), class = "data.frame", row.names = c(NA, -3))
  }

  AlleCon <- PostGeno$Freq * AM
  AllelicConcordance <- sum(AlleCon)
  #Insert table to add to.
  tmp <- cbind(ID, Genotype, GenoConcordance, AllelicConcordance, Imp_Geno_Dosage, tmp1$Freq, tmp2$Freq, tmp3$Freq)
  df_pheno_Pen <- rbind(df_pheno_Pen, tmp)
  
}

```

# Save file to prevent loss/deletion
```{r}
write.csv(df_pheno_Pen, "Pheno_Pen_AFAPLowMutation.csv")
```

##Statistical Analysis
#Pearson (Squared) Correlation Coefficient
```{r}
## Observed Genotype Only
#First add column with Obsv_geno_Dosages
df_geno_only$ObsGenoDosage <- NA
df_geno_only <- df_geno_only %>%
  transform(ObsGenoDosage = ifelse(df_geno_only$Genotype == "AA", 0, df_geno_only$ObsGenoDosage ))
df_geno_only <- df_geno_only %>%
  transform(ObsGenoDosage = ifelse(df_geno_only$Genotype == "AB", 1, df_geno_only$ObsGenoDosage ))
df_geno_only <- df_geno_only %>%
  transform(ObsGenoDosage = ifelse(df_geno_only$Genotype == "BB", 2, df_geno_only$ObsGenoDosage ))
#turn to numeric
df_geno_only$Imp_Geno_Dosage <- as.numeric(df_geno_only$Imp_Geno_Dosage)
#Correlation - (then Square for R2)
CorGenoOnlyMask <- cor.test(df_geno_only$Imp_Geno_Dosage, df_geno_only$ObsGenoDosage, method = "pearson")


## Observed Genotype and Observed Phenotype
#First add column with Obsv_geno_Dosages
df_pheno$ObsGenoDosage <- NA
df_pheno <- df_pheno %>%
  transform(ObsGenoDosage = ifelse(df_pheno$Genotype == "AA", 0, df_pheno$ObsGenoDosage ))
df_pheno <- df_pheno %>%
  transform(ObsGenoDosage = ifelse(df_pheno$Genotype == "AB", 1, df_pheno$ObsGenoDosage ))
df_pheno <- df_pheno %>%
  transform(ObsGenoDosage = ifelse(df_pheno$Genotype == "BB", 2, df_pheno$ObsGenoDosage ))
#turn to numeric
df_pheno$Imp_Geno_Dosage <- as.numeric(df_pheno$Imp_Geno_Dosage)
#Correlation - (then Square for R2)
CorPhenoMask <- cor.test(df_pheno$Imp_Geno_Dosage, df_pheno$ObsGenoDosage, method = "pearson")

## Observed Genotype and Observed Phenotype with Age-Adjusted Penetrance
#First add column with Obsv_geno_Dosages
df_pheno_Pen$ObsGenoDosage <- NA
df_pheno_Pen <- df_pheno_Pen %>%
  transform(ObsGenoDosage = ifelse(df_pheno_Pen$Genotype == "AA", 0, df_pheno_Pen$ObsGenoDosage ))
df_pheno_Pen <- df_pheno_Pen %>%
  transform(ObsGenoDosage = ifelse(df_pheno_Pen$Genotype == "AB", 1, df_pheno_Pen$ObsGenoDosage ))
df_pheno_Pen <- df_pheno_Pen %>%
  transform(ObsGenoDosage = ifelse(df_pheno_Pen$Genotype == "BB", 2, df_pheno_Pen$ObsGenoDosage ))
#turn to numeric
df_pheno_Pen$Imp_Geno_Dosage <- as.numeric(df_pheno_Pen$Imp_Geno_Dosage)
#Correlation - (then Square for R2)
CorPhenoPenMask <- cor.test(df_pheno_Pen$Imp_Geno_Dosage, df_pheno_Pen$ObsGenoDosage, method = "pearson")


```

#Testing normality
```{r}
shapiro.test(df_geno_only$AllelicConcordance)
shapiro.test(df_pheno$GenoConcordance)
shapiro.test(df_pheno_Pen$GenoConcordance)
qqnorm(df_geno_only$GenoConcordance)
qqline(df_geno_only$GenoConcordance)
qqnorm(df_pheno$GenoConcordance)
qqline(df_pheno$GenoConcordance)
qqnorm(df_pheno_Pen$GenoConcordance)
qqline(df_pheno_Pen$GenoConcordance)
```

#Running the Kruskal Test
```{r}
#Kruskal.test used to test significance between the three models (does not assume a distribution, non-parametric)
#Genotype Concordance
data_list <- list(df_geno_only$GenoConcordance, df_pheno$GenoConcordance, df_pheno_Pen$GenoConcordance)
kruskal.test(data_list)
#Allele Concordance
data_list <- list(df_geno_only$AllelicConcordance, df_pheno$AllelicConcordance, df_pheno_Pen$AllelicConcordance)
kruskal.test(data_list)
```

#Plotting barchart with error bars
```{r}
#Find the mean and standard deviation for each of the three models (as percent)
#Genotype Concordance
mean_geno_GC <- mean(df_geno_only$GenoConcordance)*100
sd_geno_GC <- sd(df_geno_only$GenoConcordance)*100

mean_pheno_GC <- mean(df_pheno$GenoConcordance)*100
sd_pheno_GC <- sd(df_pheno$GenoConcordance)*100

mean_phenoPen_GC <- mean(df_pheno_Pen$GenoConcordance)*100
sd_phenoPen_GC <- sd(df_pheno_Pen$GenoConcordance)*100

#Allele Concordance
mean_geno_AC <- mean(df_geno_only$AllelicConcordance)*100
sd_geno_AC <- sd(df_geno_only$AllelicConcordance)*100

mean_pheno_AC <- mean(df_pheno$AllelicConcordance)*100
sd_pheno_AC <- sd(df_pheno$AllelicConcordance)*100

mean_phenoPen_AC <- mean(df_pheno_Pen$AllelicConcordance)*100
sd_phenoPen_AC <- sd(df_pheno_Pen$AllelicConcordance)*100

#Combine into one dataframe
tmp <- cbind("1", "GC", mean_geno_GC, sd_geno_GC)
tmp2 <- cbind("1", "AC", mean_geno_AC, sd_geno_AC)

tmp3 <- cbind("2", "GC", mean_pheno_GC, sd_pheno_GC)
tmp4 <- cbind("2", "AC", mean_pheno_AC, sd_pheno_AC)

tmp5 <- cbind("3", "GC", mean_phenoPen_GC, sd_phenoPen_GC)
tmp6 <- cbind("3", "AC", mean_phenoPen_AC, sd_phenoPen_AC)


summary <- data.frame(rbind(tmp, tmp2, tmp3, tmp4, tmp5, tmp6))
summary$mean_geno_GC <- as.numeric(summary$mean_geno_GC)
summary$sd_geno_GC <- as.numeric(summary$sd_geno_GC)

ggplot(summary, aes(V1, mean_geno_GC, fill = V2)) +geom_bar(stat = "identity", color = "black", position = position_dodge())+
  geom_errorbar(aes(ymin = mean_geno_GC-sd_geno_GC, ymax = mean_geno_GC + sd_geno_GC), width =.2, position=position_dodge(.9)) +
  xlab("Model") + scale_x_discrete(labels = c("1", "2", "3")) +
  scale_fill_manual(values = c("#CC79A7","#7205FF"), labels = c("Allele Concordance", "Genotype Concordance"), name = "") +
  ylab("Concordance (%)")
```


#Plotting histograms for Genotype and Allelic Concordance
```{r}
df_geno_only %>% ggplot() + aes(GenoConcordance)+ geom_histogram()
df_pheno %>% ggplot() + aes(GenoConcordance)+ geom_histogram()
df_pheno_Pen %>% ggplot() + aes(GenoConcordance)+ geom_histogram()

df_geno_only$GenoConcordance <- as.numeric(df_geno_only$GenoConcordance)
df_pheno$GenoConcordance <- as.numeric(df_pheno$GenoConcordance)
df_pheno_Pen$GenoConcordance <- as.numeric(df_pheno_Pen$GenoConcordance)

df_geno_only$GenoConcordance100 <- df_geno_only$GenoConcordance*100
df_pheno$GenoConcordance100 <- df_pheno$GenoConcordance*100
df_pheno_Pen$GenoConcordance100 <- df_pheno_Pen$GenoConcordance*100

a <- ggplot(df_geno_only) + aes(GenoConcordance100, fill = Genotype) + geom_histogram(bins = 40, na.rm = TRUE) + scale_fill_manual(values = c("#009E73", "#0072B2", "#D55E00"), name = "Genotype", labels = c("AA", "AB", "BB")) + xlab("Genotype Concordance (%)")+ xlim(-1, 101) +ylim(0, 20)
b <- ggplot(df_pheno) + aes(GenoConcordance100, fill = Genotype) + geom_histogram(bins = 40, na.rm = TRUE) + scale_fill_manual(values = c("#009E73", "#0072B2", "#D55E00"), name = "Genotype", labels = c("AA", "AB", "BB")) + xlab("Genotype Concordance (%)") + xlim(-1, 101) +ylim(0, 20)
c <- ggplot(df_pheno_Pen) + aes(GenoConcordance100, fill = Genotype) + geom_histogram(bins = 40, na.rm = TRUE) + scale_fill_manual(values = c("#009E73", "#0072B2", "#D55E00"), name = "Genotype", labels = c("AA", "AB", "BB")) + xlab("Genotype Concordance (%)") + xlim(-1, 101) +ylim(0, 20)


ggarrange(a, b, c, labels = c("1", "2", "3"), common.legend = TRUE, legend = "right", nrow = 3, ncol = 1)

#Allele Concordance
df_geno_only$AllelicConcordance <- as.numeric(df_geno_only$AllelicConcordance)
df_pheno$AllelicConcordance <- as.numeric(df_pheno$AllelicConcordance)
df_pheno_Pen$AllelicConcordance <- as.numeric(df_pheno_Pen$AllelicConcordance)
df_geno_only$AllelicConcordance100 <- df_geno_only$AllelicConcordance*100
df_pheno$AllelicConcordance100 <- df_pheno$AllelicConcordance*100
df_pheno_Pen$AllelicConcordance100 <- df_pheno_Pen$AllelicConcordance*100

a <- ggplot(df_geno_only) + aes(AllelicConcordance100, fill = Genotype) + geom_histogram(bins = 40, na.rm = TRUE) + scale_fill_manual(values = c("#009E73", "#0072B2", "#D55E00"), name = "Genotype", labels = c("AA", "AB", "BB")) + xlab("Allele Concordance (%)") + xlim(-1, 101)
b <- ggplot(df_pheno) + aes(AllelicConcordance100, fill = Genotype) + geom_histogram(bins = 40, na.rm = TRUE) + scale_fill_manual(values = c("#009E73", "#0072B2", "#D55E00"), name = "Genotype", labels = c("AA", "AB", "BB")) + xlab("Allele Concordance (%)") + xlim(-1, 101)
c <- ggplot(df_pheno_Pen) + aes(AllelicConcordance100, fill = Genotype) + geom_histogram(bins = 40, na.rm = TRUE) + scale_fill_manual(values = c("#009E73", "#0072B2", "#D55E00"), name = "Genotype", labels = c("AA", "AB", "BB")) + xlab("Allelic Concordance (%)") + xlim(-1, 101)


ggarrange(a, b, c, labels = c("1", "2", "3"), common.legend = TRUE, legend = "right", nrow = 3, ncol = 1)

```

