---
title: "Sub-Pedigree Analysis"
output: html_document
date: '2023-05-10'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Clean session
rm(list = ls())
library(tidyverse)
library(ggplot2)
library(ggforce)
library(reshape2)
library(ggpubr)

# Install AlleleFetch
if (FALSE) {
  setRepositories() # choose CRAN and BioC
  install.packages(pkg = c("gRain", "devtools"), dep = TRUE)
  # PR that fixed the repeatPattern() data issue
  # https://github.com/hojsgaard/gRain/pull/9
  devtools::install_github(repo = "https://github.com/gregorgorjanc/gRain")
}
# Load package
library(package = "gRain")

# User defined functions

source(file = "AlleleFetch.R")
```

#Insert the Sub-Pedigree
```{r}
GenoScale <- read.csv("/Users/roscraddock/Documents/University of Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/AlleleFetch Inputs/PBGV_Geno_2009-2014.csv")
GenoScale <- GenoScale %>%
  select(!X)
PhenoScale <- read.csv("/Users/roscraddock/Documents/University of Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/AlleleFetch Inputs/PBGV_Pheno_2009-2014.csv")
PhenoScale <- PhenoScale %>%
  select(!X)
PedScale <- read.csv("/Users/roscraddock/Documents/University of Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/AlleleFetch Inputs/PBGV_Ped_2009-2014.csv")
PedScale <- PedScale %>%
  select(!X)
PBGV_POAG_Sub <- read.csv("/Users/roscraddock/Documents/University\ of\ Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/PBGV_Sub_2009-2014.csv")
PBGV_POAG_Sub <- PBGV_POAG_Sub %>%
  select(!X)
```

#Pedigree visualisation - Sub pedigree
```{r}
library(visPedigree)

SubPedDiag <- PBGV_POAG_Sub %>%
  select(IRN,IRN_sire,IRN_dam,Sex) %>%
  transform(Sex=ifelse(Sex =="Dog", "male", "female"))
SubPedDiag <- tidyped(SubPedDiag)
visped(SubPedDiag, compact = TRUE, cex = 0.83, file = "SubPedDiag.pdf")
```

#Setting Up for AlleleFetch
```{r}
#Setting up the genotype and phenotype states
testGenotypeState <- c("AA", "AB", "BB")
testPhenotypeState <- c("OK", "NOK")

# Prior allele frequency in (founder) population - #0.309474 taken from AlphaPeel
PrB <- 0.309474 
PrA <- 1-PrB


# Prior genotype frequencies in (founder) population
# according to the Hardy-Weinberg's expectation
testGenoPT <- c(PrA*PrA, 2*PrA*PrB, PrB*PrB)

# Genotype frequencies in individual given the genotype of parents according to Mendel's law
# and accounting for mutation (via e)
e <- 0.00000001
gM_AA_AA <- c(1-2*e+e^2, 2*e-2*e^2, e^2)
gM_AA_AB <- c(1/2-1/2*e, 1/2, 1/2*e)
gM_AA_BB <- c(e-e^2, 1-2*e+2*e^2, e-e^2)
gM_AB_AA <- c(1/2-1/2*e, 1/2, 1/2*e)
gM_AB_AB <- c(1/4, 1/2, 1/4)
gM_AB_BB <- c(1/2*e, 1/2, 1/2-1/2*e)
gM_BB_AA <- c(e-e^2, 1-2*e+2*e^2, e-e^2)
gM_BB_AB <- c(1/2*e, 1/2, 1/2-1/2*e)
gM_BB_BB <- c(e^2, 2*e-2*e^2, 1-2*e+e^2)

testMendelianCPT <- matrix(nrow=3, ncol=9)
testMendelianCPT[, 1] <- gM_AA_AA
testMendelianCPT[, 2] <- gM_AA_AB
testMendelianCPT[, 3] <- gM_AA_BB
testMendelianCPT[, 4] <- gM_AB_AA
testMendelianCPT[, 5] <- gM_AB_AB
testMendelianCPT[, 6] <- gM_AB_BB
testMendelianCPT[, 7] <- gM_BB_AA
testMendelianCPT[, 8] <- gM_BB_AB
testMendelianCPT[, 9] <- gM_BB_BB

# Genotype frequencies in individual given the genotype of one parent (Mendel's law with mutation)
# and base population (Hardy-Weinberg's law)
gM_AA_unk <- c(PrA-PrA*e, PrB-PrB*e+PrA*e, PrB*e)
gM_AB_unk <- c(1/2*PrA, 1/2*(PrA+PrB), 1/2*PrB)
gM_BB_unk <- c(PrA*e, PrA-PrA*e+PrB*e, PrB-PrB*e)

testMendelianCPTOnlyFather <- matrix(nrow=3, ncol= 3)
testMendelianCPTOnlyFather[,1] <- gM_AA_unk
testMendelianCPTOnlyFather[,2] <- gM_AB_unk
testMendelianCPTOnlyFather[,3] <- gM_BB_unk

testMendelianCPTOnlyMother <- testMendelianCPTOnlyFather

# Set of phenotype values
testPhenotypeState <- c("OK", "NOK")

# Phenotype frequences in individual given the genotype of the individual - phenotype penetrance
e <- 0.01
testPhenoPenetranceCPT <- matrix(nrow=2, ncol=3)
#                               OK   NOK Pheno
testPhenoPenetranceCPT[,1] <- c(1-e, 0+e) # AA True geno
testPhenoPenetranceCPT[,2] <- c(1-e, 0+e) # AB
testPhenoPenetranceCPT[,3] <- c(0+e, 1-e) # BB   

# Genotype frequences in individual given the genotype of the individual - genotype penetrance
e <- 0.1
testGenoPenetranceCPT <- matrix(nrow=3, ncol=3)
#                             AA       AB     BB Observed Geno
testGenoPenetranceCPT[,1] <-c(1-e-e^2, e,     e^2)     # AA True Geno
testGenoPenetranceCPT[,2] <-c(e,       1-2*e, e)       # AB
testGenoPenetranceCPT[,3] <-c(e^2,     e,     1-e-e^2) # BB
```

#Running through AlleleFetch without any masking - Genotype only
```{r}
out <- AlleleFetch(pedigreeData = PedScale,
                   genotypeState = testGenotypeState,
                   genotypeData = GenoScale,
                   phenotypeState = testPhenotypeState,
                   genoPT = testGenoPT,
                   mendelianCPT = testMendelianCPT,
                   mendelianCPTOnlyFather = testMendelianCPTOnlyFather,
                   mendelianCPTOnlyMother = testMendelianCPTOnlyMother,
                   genoPenetranceCPT = testGenoPenetranceCPT,
                   phenoPenetranceCPT = testPhenoPenetranceCPT)
#Save results into dataframe
Result <- out$result
#Save results
write.csv(Result, "/Users/roscraddock/Documents/University of Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/AlleleFetch Inputs/PBGV_Result_2009-2014.csv" )

```

#Running through AlleleFetch without any masking - Genotype and Phenotype
```{r}
out <- AlleleFetch(pedigreeData = PedScale,
                   genotypeState = testGenotypeState,
                   genotypeData = GenoScale,
                   phenotypeState = testPhenotypeState,
                   phenotypeData = PhenoScale,
                   genoPT = testGenoPT,
                   mendelianCPT = testMendelianCPT,
                   mendelianCPTOnlyFather = testMendelianCPTOnlyFather,
                   mendelianCPTOnlyMother = testMendelianCPTOnlyMother,
                   genoPenetranceCPT = testGenoPenetranceCPT,
                   phenoPenetranceCPT = testPhenoPenetranceCPT)
#Save results into dataframe
ResultPhen <- out$result
#Save results
write.csv(ResultPhen, "/Users/roscraddock/Documents/University of Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/AlleleFetch Inputs/PBGV_ResultPheno_2009-2014.csv" )
```

#Running through for age adjusted penetrance
```{r}

#Additional phenotype function for age at eye test under 4.
testPhenoPenetranceCPT_Before <- matrix(nrow=2, ncol=3)
 #                           OK<4 NOK<4 Pheno
 testPhenoPenetranceCPT_Before[,1] <- c(1-e, 0+e) # AA True geno
 testPhenoPenetranceCPT_Before[,2] <- c(1-e, 0+e) # AB
 testPhenoPenetranceCPT_Before[,3] <- c(1-e, 0+e) # BB
 
 #Additional dataset for the age at eye testing
PBGV_POAG_Sub <- read.csv("/Users/roscraddock/Documents/University\ of\ Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/PBGV_Sub_2009-2014.csv")

AgePhenoScale <- PBGV_POAG_Sub %>%
  select(IRN, DOB, EyeTest_Date)
AgePhenoScale <- AgePhenoScale %>%
  filter(!EyeTest_Date == "")
AgePhenoScale$DOB <- as.Date(AgePhenoScale$DOB)
AgePhenoScale$EyeTest_Date <- as.Date(AgePhenoScale$EyeTest_Date)

AgePhenoScale$Age <- AgePhenoScale$EyeTest_Date - AgePhenoScale$DOB
AgePhenoScale$Age <- as.numeric(AgePhenoScale$Age) %/% 365.25

AgePhenoScale <- AgePhenoScale %>%
  select(IRN, Age) %>%
  rename(IID = "IRN")

#Run AlleleFetch
  out <- AlleleFetch(pedigreeData = PedScale,
                   genotypeState = testGenotypeState,
                   genotypeData = GenoScale,
                   phenotypeState = testPhenotypeState,
                   phenotypeData = PhenoScale,
                   genoPT = testGenoPT,
                   mendelianCPT = testMendelianCPT,
                   mendelianCPTOnlyFather = testMendelianCPTOnlyFather,
                   mendelianCPTOnlyMother = testMendelianCPTOnlyMother,
                   genoPenetranceCPT = testGenoPenetranceCPT,
                   phenoPenetranceCPT = testPhenoPenetranceCPT,
                   phenoPenetranceCPT_Before = testPhenoPenetranceCPT_Before,
                   agePhenotype = AgePhenoScale)
  #Save results into dataframe
  ResultPhenAge <- out$result
  write.csv(ResultPhenAge, "/Users/roscraddock/Documents/University of Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/AlleleFetch Inputs/PBGV_ResultPhenoAge_2009-2014.csv" )
```

#Plotting the Trends for AlphaPeel, and the three models in AlleleFetch
(1)When ran through AlphaPeel
```{r}
AP_Sub_Data <- read.csv("/Users/roscraddock/Documents/University\ of\ Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/Alpha_Peel_2009-2014.csv")
AP_Sub_Result <- read.csv("/Users/roscraddock/Documents/Agriculture Uni/Year 3/NERC Placement/AlphaPeel Practice/AlphaPeel-master/AlphaPeel/outputs/SingleLocus2009-2014.haps", header = FALSE, sep = " ", col.names = c("IID", "Genotype_Probabilities"))

#Add the genotype states/values
x <- c("AA", "BA", "AB", "BB")
x <- data.frame(x)
x <- x %>%
  rename(Geno_State = "x")
test <- x
ID <- ((as.numeric(count(AP_Sub_Result)))/4)-1

for(i in 1:ID) {
test <- rbind(test, x)
}

AP_Sub_Result <- cbind(AP_Sub_Result, test)

#Move to wide format
AP_Sub_Result <- AP_Sub_Result %>%
    pivot_wider(names_from = Geno_State,
                values_from = Genotype_Probabilities)

#Remove "mother of.." and "father of..."
AP_Sub_Result$IID <- as.numeric(AP_Sub_Result$IID)
AP_Sub_Result <- AP_Sub_Result %>%
  filter(!is.na(IID))

DOB <-AP_Sub_Data %>%
  select(IID, DOB)

AP_Sub_Result <- merge(AP_Sub_Result, DOB, by="IID")

#Since the order of alleles is not relevant in this example, add columns Aa and aA together
AP_Sub_Result$Het <- AP_Sub_Result$BA + AP_Sub_Result$AB
AP_Sub_Result <- AP_Sub_Result %>%
  select(IID, AA, Het, BB, DOB) %>%
  rename(AB = "Het")

#Remove individuals with no recorded date of birth
AP_Sub_Result <- AP_Sub_Result %>% 
  filter(!is.na(DOB))

# Change DOB format to date
AP_Sub_Result$DOB <- as.Date(AP_Sub_Result$DOB)

# Add year and month columns
AP_Sub_Result$year <- format(AP_Sub_Result$DOB, "%Y")
AP_Sub_Result$month <- format(AP_Sub_Result$DOB, "%m")
AP_Sub_Result$year_month <- format(AP_Sub_Result$DOB, "%Y-%m")
str(AP_Sub_Result)

#By year
tmp <- AP_Sub_Result %>%
  group_by(year) %>%
  summarise(mean(AA))
tmp2 <- AP_Sub_Result %>%
  group_by(year) %>%
  summarise(mean(AB))
tmp3 <- AP_Sub_Result %>%
  group_by(year) %>%
  summarise(mean(BB))

AP_Sub_Result_Plot <- cbind(tmp, tmp2, tmp3)

#Reduce to only one column for year
AP_Sub_Result_Plot <- AP_Sub_Result_Plot[,c(1,2,4,6),]

#Move back into the tall format
AP_Sub_Result_Plot <- melt(AP_Sub_Result_Plot, id.vars="year", variable.name = "State" )

# Plotting the Mean frequency of states per year (1972 to January 2023)
AP <- AP_Sub_Result_Plot %>% ggplot(aes(year, value, group=State, col=State))+
  geom_line(size=1) +
  xlab("Year of Birth") +
  ylab("Mean Frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  scale_colour_manual(values = c("#009E73", "#0072B2", "#D55E00"), name = "Genotype", labels = c("AA", "AB", "BB"))+
  geom_vline(xintercept = "1995", linetype = "dotted") +
  geom_vline(xintercept = "2015", linetype = "dashed", colour = "#D35FB7")

```

(2) when ran through AlleleFetch with genotypes
```{r}
AF_Sub_Data <- read.csv("/Users/roscraddock/Documents/University\ of\ Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/PBGV_Sub_2009-2014.csv")
AF_Sub_Result <- read.csv("/Users/roscraddock/Documents/University of Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/AlleleFetch Inputs/PBGV_Result_2009-2014.csv")
AF_Sub_Result <- AF_Sub_Result %>%
  select(!X)

#Widen AF_Sub_Result
AF_Sub_Result <- AF_Sub_Result %>%
  pivot_wider(names_from = c(State, Var), 
              values_from = Freq)
#Merge together
AF_Sub_Trend <- merge(AF_Sub_Result, AF_Sub_Data, by.x = "IID", by.y = "IRN")

#Reduce to the IID, Genotypes, and DOB
AF_Sub_Trend <- AF_Sub_Trend %>%
  select(IID, AA_geno, AB_geno, BB_geno, OK_pheno, NOK_pheno, DOB)

#Remove where DOB is NA - 800 removed?
AF_Sub_Trend <- AF_Sub_Trend[!is.na(AF_Sub_Trend$DOB),]
#Set DOB as date
AF_Sub_Trend$DOB <- as.Date(AF_Sub_Trend$DOB)
#Separate year into table
AF_Sub_Trend$year <- format(AF_Sub_Trend$DOB, "%Y")

#By year
tmp <- AF_Sub_Trend %>%
  group_by(year) %>%
  summarise(mean(AA_geno))
tmp2 <- AF_Sub_Trend %>%
  group_by(year) %>%
  summarise(mean(AB_geno))
tmp3 <- AF_Sub_Trend %>%
  group_by(year) %>%
  summarise(mean(BB_geno))
tmp4 <- AF_Sub_Trend %>%
  group_by(year) %>%
  summarise(mean(OK_pheno))
tmp5 <- AF_Sub_Trend %>%
  group_by(year) %>%
  summarise(mean(NOK_pheno))

AF_Sub_Trend_Plot <- cbind(tmp,tmp2,tmp3,tmp4,tmp5)
AF_Sub_Trend_Plot <- AF_Sub_Trend_Plot[,c(1,2,4,6, 8, 10),]

#Move back into the tall format
AF_Sub_Trend_Plot <- melt(AF_Sub_Trend_Plot, id.vars="year", variable.name = "State" )

# Plotting the Mean frequency of states per year (2009 to January 2014)
AFGeno <- AF_Sub_Trend_Plot %>% ggplot(aes(year, value, group=State, col=State))+
  geom_line(size=1) +
  xlab("Year of Birth") +
  ylab("Mean Frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  scale_colour_manual(values = c("#009E73", "#0072B2", "#D55E00", "#E69F00", "#7C788A"), name = "Genotype", labels = c("AA", "AB", "BB", "Unaffected", "Affected"))+
  geom_vline(xintercept = "1995", linetype = "dotted") +
  geom_vline(xintercept = "2015", linetype = "dashed", colour = "#D35FB7")

```

(3) when ran through AlleleFetch with genotypes and phenotypes
```{r}
AF_Sub_Data <- read.csv("/Users/roscraddock/Documents/University\ of\ Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/PBGV_Sub_2009-2014.csv")
AF_Sub_ResultPheno <- read.csv("/Users/roscraddock/Documents/University of Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/AlleleFetch Inputs/PBGV_ResultPheno_2009-2014.csv")
AF_Sub_ResultPheno <- AF_Sub_ResultPheno %>%
  select(!X)

#Widen AF_Sub_Result
AF_Sub_ResultPheno <- AF_Sub_ResultPheno %>%
  pivot_wider(names_from = c(State, Var), 
              values_from = Freq)
#Merge together
AF_SubPheno_Trend <- merge(AF_Sub_ResultPheno, AF_Sub_Data, by.x = "IID", by.y = "IRN")

#Reduce to the IID, Genotypes, and DOB
AF_SubPheno_Trend <- AF_SubPheno_Trend %>%
  select(IID, AA_geno, AB_geno, BB_geno, OK_pheno, NOK_pheno, DOB)

#Remove where DOB is NA - 800 removed?
AF_SubPheno_Trend <- AF_SubPheno_Trend[!is.na(AF_SubPheno_Trend$DOB),]
#Set DOB as date
AF_SubPheno_Trend$DOB <- as.Date(AF_SubPheno_Trend$DOB)
#Separate year into table
AF_SubPheno_Trend$year <- format(AF_SubPheno_Trend$DOB, "%Y")

#By year
tmp <- AF_SubPheno_Trend %>%
  group_by(year) %>%
  summarise(mean(AA_geno))
tmp2 <- AF_SubPheno_Trend %>%
  group_by(year) %>%
  summarise(mean(AB_geno))
tmp3 <- AF_SubPheno_Trend %>%
  group_by(year) %>%
  summarise(mean(BB_geno))
tmp4 <- AF_SubPheno_Trend %>%
  group_by(year) %>%
  summarise(mean(OK_pheno))
tmp5 <- AF_SubPheno_Trend %>%
  group_by(year) %>%
  summarise(mean(NOK_pheno))

AF_SubPheno_Trend_Plot <- cbind(tmp,tmp2,tmp3,tmp4,tmp5)
AF_SubPheno_Trend_Plot <- AF_SubPheno_Trend_Plot[,c(1,2,4,6, 8, 10),]

#Move back into the tall format
AF_SubPheno_Trend_Plot <- melt(AF_SubPheno_Trend_Plot, id.vars="year", variable.name = "State" )

# Plotting the Mean frequency of states per year (2009 to January 2014)
AFPheno <- AF_SubPheno_Trend_Plot %>% ggplot(aes(year, value, group=State, col=State))+
  geom_line(size=1) +
  xlab("Year of Birth") +
  ylab("Mean Frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  scale_colour_manual(values = c("#009E73", "#0072B2", "#D55E00", "#E69F00", "#7C788A"), name = "Genotype", labels = c("AA", "AB", "BB", "Unaffected", "Affected"))+
  geom_vline(xintercept = "1995", linetype = "dotted") +
  geom_vline(xintercept = "2015", linetype = "dashed", colour = "#D35FB7")

```

(4) When run through AlleleFetch with DNA tests, eye tests, and age-adjusted penetrance.
```{r}
AF_Sub_Data <- read.csv("/Users/roscraddock/Documents/University\ of\ Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/PBGV_Sub_2009-2014.csv")
AF_Sub_ResultAge <- read.csv("/Users/roscraddock/Documents/University of Edinburgh/HighlanderLab - AlleleFetch_data/Clean Pedigrees/PBGV/AlleleFetch Inputs/PBGV_ResultPhenoAge_2009-2014.csv")
AF_Sub_ResultAge <- AF_Sub_ResultAge %>%
  select(!X)

#Widen AF_Sub_Result
AF_Sub_ResultAge <- AF_Sub_ResultAge %>%
  pivot_wider(names_from = c(State, Var), 
              values_from = Freq)
#Merge together
AF_Sub_TrendAge <- merge(AF_Sub_ResultAge, AF_Sub_Data, by.x = "IID", by.y = "IRN")

#Reduce to the IID, Genotypes, and DOB
AF_Sub_TrendAge <- AF_Sub_TrendAge %>%
  select(IID, AA_geno, AB_geno, BB_geno, OK_pheno, NOK_pheno, DOB)

#Remove where DOB is NA - 800 removed?
AF_Sub_TrendAge <- AF_Sub_TrendAge[!is.na(AF_Sub_TrendAge$DOB),]
#Set DOB as date
AF_Sub_TrendAge$DOB <- as.Date(AF_Sub_TrendAge$DOB)
#Separate year into table
AF_Sub_TrendAge$year <- format(AF_Sub_TrendAge$DOB, "%Y")

#By year
tmp <- AF_Sub_TrendAge %>%
  group_by(year) %>%
  summarise(mean(AA_geno))
tmp2 <- AF_Sub_TrendAge %>%
  group_by(year) %>%
  summarise(mean(AB_geno))
tmp3 <- AF_Sub_TrendAge %>%
  group_by(year) %>%
  summarise(mean(BB_geno))
tmp4 <- AF_Sub_TrendAge %>%
  group_by(year) %>%
  summarise(mean(OK_pheno))
tmp5 <- AF_Sub_TrendAge %>%
  group_by(year) %>%
  summarise(mean(NOK_pheno))

AF_Sub_TrendAge_Plot <- cbind(tmp,tmp2,tmp3,tmp4,tmp5)
AF_Sub_TrendAge_Plot <- AF_Sub_TrendAge_Plot[,c(1,2,4,6, 8, 10),]

#Move back into the tall format
AF_Sub_TrendAge_Plot <- melt(AF_Sub_TrendAge_Plot, id.vars="year", variable.name = "State" )

# Plotting the Mean frequency of states per year (2009 to January 2014)
AFPhenoAge <- AF_Sub_TrendAge_Plot %>% ggplot(aes(year, value, group=State, col=State))+
  geom_line(size=1) +
  xlab("Year of Birth") +
  ylab("Mean Frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  scale_colour_manual(values = c("#009E73", "#0072B2", "#D55E00", "#E69F00", "#7C788A"), name = "Genotype", labels = c("AA", "AB", "BB", "Unaffected", "Affected"))+
  geom_vline(xintercept = "1995", linetype = "dotted") +
  geom_vline(xintercept = "2015", linetype = "dashed", colour = "#D35FB7")

```

Plotting all four plots together
```{r}
ggarrange(AP, AFGeno, AFPheno, AFPhenoAge, labels = c("a", "b", "c", "d"), ncol= 2, nrow = 2, common.legend = TRUE, legend = "right")

#Plotting with the section taken from the full pedigree

```


#Plotting the change in the genotype probabilities for the 11 dogs eye tested only.
```{r}
test <- PBGV_POAG_Sub %>%
  filter(!EyeTest_Result == "") %>%
  filter(DNA_result == "" | is.na(DNA_result))

tmp <- Result %>%
  filter(IID == 8685312) %>%
  filter(!Var == "pheno") %>%
  group_by(IID)

tmp <- Result$IID %in% test$IRN
OPGResults <- Result[tmp,]
OPGResults <- OPGResults %>%
  filter(!Var == "pheno")

tmp <- ResultPhen$IID %in% test$IRN
OPGPResults <- ResultPhen[tmp,]
OPGPResults <- OPGPResults %>%
  filter(!Var == "pheno")

tmp <- ResultPhenAge$IID %in% test$IRN
OPAGP <- ResultPhenAge[tmp,]
OPAGP <- OPGPResults %>%
  filter(!Var == "pheno")

OPGResults$IID <- as.character(OPGResults$IID)

OPGPlot <- ggplot(OPGResults, aes(IID, Freq, fill = State)) + geom_col(colour = "black", position = "fill") +
  scale_fill_manual(values = c("#009E73", "#0072B2", "#D55E00"), name = "Genotype") + ylab("Frequency") + xlab("Dog") + scale_x_discrete(labels = c(1,2,3,4,5,6,7,8,9,10,11))

OPGPResults$IID <- as.character(OPGPResults$IID)
OPGPPlot <- ggplot(OPGPResults, aes(IID, Freq, fill = State)) + geom_col(colour = "black", position = "fill") +
  scale_fill_manual(values = c("#009E73", "#0072B2", "#D55E00"), name = "Genotype") + ylab("Frequency") + xlab("Dog") + scale_x_discrete(labels = c(1,2,3,4,5,6,7,8,9,10,11))

ggarrange(OPGPlot, OPGPPlot, labels = c("a", "b"), ncol= 1, nrow = 2, common.legend = TRUE, legend = "right")

OPAGP$IID <- as.character(OPAGP$IID)

#Checking age-adjusted model is the same.
OPAGPPlot <- ggplot(OPAGP, aes(IID, Freq, fill = State)) + geom_col(colour = "black", position = "fill") +
  scale_fill_manual(values = c("#009E73", "#0072B2", "#D55E00"), name = "Genotype") + ylab("Frequency") + xlab("Dog") + scale_x_discrete(labels = c(1,2,3,4,5,6,7,8,9,10,11))

ggarrange(OPGPlot, OPGPPlot, OPAGPPlot, labels = c("a", "b", "c"), ncol= 1, nrow = 3, common.legend = TRUE, legend = "right")
```


#Example for the identification of At-Risk Dogs
```{r}
#Mean Minor Allele Frequency per year
#DNA test only
#Widen AF_Sub_Trend_Plot
AF_Sub_MAF <- AF_Sub_Trend_Plot %>%
  pivot_wider(names_from = State, 
              values_from = value)
AF_Sub_MAF$MAF <- AF_Sub_MAF$`mean(BB_geno)`+0.5*AF_Sub_MAF$`mean(AB_geno)`
#DNA test and phenotype
AF_SubPheno_MAF <- AF_SubPheno_Trend_Plot %>%
  pivot_wider(names_from = State, 
              values_from = value)
AF_SubPheno_MAF$MAF <- AF_SubPheno_MAF$`mean(BB_geno)`+0.5*AF_SubPheno_MAF$`mean(AB_geno)`
#DNA test, eye test, and age-adjusted penetrance
AF_SubAge_MAF <- AF_Sub_TrendAge_Plot %>%
  pivot_wider(names_from = State, 
              values_from = value)
AF_SubAge_MAF$MAF <- AF_SubAge_MAF$`mean(BB_geno)`+0.5*AF_SubAge_MAF$`mean(AB_geno)`


#2013 - BB, DNA test only
At_Risk_2013 <- AF_Sub_Trend[AF_Sub_Trend$DOB > "2012-12-31",]
At_Risk_2013 <- At_Risk_2013[At_Risk_2013$DOB < "2014-01-01",]
At_Risk_2013 <- At_Risk_2013[!is.na(At_Risk_2013$IID),]
At_Risk_2013 <- At_Risk_2013 %>%
  filter(BB_geno >0.0615)

x <- AF_Sub_Data$IRN %in% At_Risk_2013$IID 
y <- AF_Sub_Data[x,]
UnknownDNA <- y %>%
  filter(DNA_result == "" | DNA_result == NA)

#2013 - AB, DNA test only
At_Risk_2013 <- AF_Sub_Trend[AF_Sub_Trend$DOB > "2012-12-31",]
At_Risk_2013 <- At_Risk_2013[At_Risk_2013$DOB < "2014-01-01",]
At_Risk_2013 <- At_Risk_2013[!is.na(At_Risk_2013$IID),]
At_Risk_2013 <- At_Risk_2013 %>%
  filter(AB_geno >0.373)

x <- AF_Sub_Data$IRN %in% At_Risk_2013$IID 
y <- AF_Sub_Data[x,]
UnknownDNA <- y %>%
  filter(DNA_result == "" | DNA_result == NA)

#2013 - BB, DNA test and eye test
At_Risk_2013 <- AF_SubPheno_Trend[AF_SubPheno_Trend$DOB > "2012-12-31",]
At_Risk_2013 <- At_Risk_2013[At_Risk_2013$DOB < "2014-01-01",]
At_Risk_2013 <- At_Risk_2013[!is.na(At_Risk_2013$IID),]
At_Risk_2013 <- At_Risk_2013 %>%
  filter(BB_geno >0.0538)

x <- AF_Sub_Data$IRN %in% At_Risk_2013$IID 
y <- AF_Sub_Data[x,]
UnknownDNA <- y %>%
  filter(DNA_result == "" | DNA_result == NA)

#2013 - AB, DNA test and eye test
At_Risk_2013 <- AF_SubPheno_Trend[AF_SubPheno_Trend$DOB > "2012-12-31",]
At_Risk_2013 <- At_Risk_2013[At_Risk_2013$DOB < "2014-01-01",]
At_Risk_2013 <- At_Risk_2013[!is.na(At_Risk_2013$IID),]
At_Risk_2013 <- At_Risk_2013 %>%
  filter(AB_geno >0.356)

x <- AF_Sub_Data$IRN %in% At_Risk_2013$IID 
y <- AF_Sub_Data[x,]
UnknownDNA <- y %>%
  filter(DNA_result == "" | DNA_result == NA)

#2013 - BB, DNA test and eye test with age-adjusted phenotypic penetrance
At_Risk_2013 <- AF_Sub_TrendAge[AF_Sub_TrendAge$DOB > "2012-12-31",]
At_Risk_2013 <- At_Risk_2013[At_Risk_2013$DOB < "2014-01-01",]
At_Risk_2013 <- At_Risk_2013[!is.na(At_Risk_2013$IID),]
At_Risk_2013 <- At_Risk_2013 %>%
  filter(BB_geno >0.0586)

x <- AF_Sub_Data$IRN %in% At_Risk_2013$IID 
y <- AF_Sub_Data[x,]
UnknownDNA <- y %>%
  filter(DNA_result == "" | DNA_result == NA)

#2013 - AB, DNA test and eye test with age-adjusted phenotypic penetrance
At_Risk_2013 <- AF_Sub_TrendAge[AF_Sub_TrendAge$DOB > "2012-12-31",]
At_Risk_2013 <- At_Risk_2013[At_Risk_2013$DOB < "2014-01-01",]
At_Risk_2013 <- At_Risk_2013[!is.na(At_Risk_2013$IID),]
At_Risk_2013 <- At_Risk_2013 %>%
  filter(AB_geno >0.367)

x <- AF_Sub_Data$IRN %in% At_Risk_2013$IID 
y <- AF_Sub_Data[x,]
UnknownDNA <- y %>%
  filter(DNA_result == "" | DNA_result == NA)

#2012 - BB, DNA test only
At_Risk_2012 <- AF_Sub_Trend[AF_Sub_Trend$DOB > "2011-12-31",]
At_Risk_2012 <- At_Risk_2012[At_Risk_2012$DOB < "2013-01-01",]
At_Risk_2012 <- At_Risk_2012[!is.na(At_Risk_2012$IID),]
At_Risk_2012 <- At_Risk_2012 %>%
  filter(BB_geno >0.0778)

x <- AF_Sub_Data$IRN %in% At_Risk_2012$IID 
y <- AF_Sub_Data[x,]
UnknownDNA <- y %>%
  filter(DNA_result == "" | DNA_result == NA)

#2012 - AB, DNA test only
At_Risk_2012 <- AF_Sub_Trend[AF_Sub_Trend$DOB > "2011-12-31",]
At_Risk_2012 <- At_Risk_2012[At_Risk_2012$DOB < "2013-01-01",]
At_Risk_2012 <- At_Risk_2012[!is.na(At_Risk_2012$IID),]
At_Risk_2012 <- At_Risk_2012 %>%
  filter(AB_geno >0.402)

x <- AF_Sub_Data$IRN %in% At_Risk_2012$IID 
y <- AF_Sub_Data[x,]
UnknownDNA <- y %>%
  filter(DNA_result == "" | DNA_result == NA)

#2012 - BB, DNA test and eye test
At_Risk_2012 <- AF_SubPheno_Trend[AF_SubPheno_Trend$DOB > "2011-12-31",]
At_Risk_2012 <- At_Risk_2012[At_Risk_2012$DOB < "2013-01-01",]
At_Risk_2012 <- At_Risk_2012[!is.na(At_Risk_2012$IID),]
At_Risk_2012 <- At_Risk_2012 %>%
  filter(BB_geno >0.0687)

x <- AF_Sub_Data$IRN %in% At_Risk_2012$IID 
y <- AF_Sub_Data[x,]
UnknownDNA <- y %>%
  filter(DNA_result == "" | DNA_result == NA)

#2012 - AB, DNA test and eye test
At_Risk_2012 <- AF_SubPheno_Trend[AF_SubPheno_Trend$DOB > "2011-12-31",]
At_Risk_2012 <- At_Risk_2012[At_Risk_2012$DOB < "2013-01-01",]
At_Risk_2012 <- At_Risk_2012[!is.na(At_Risk_2012$IID),]
At_Risk_2012 <- At_Risk_2012 %>%
  filter(AB_geno >0.387)

x <- AF_Sub_Data$IRN %in% At_Risk_2012$IID 
y <- AF_Sub_Data[x,]
UnknownDNA <- y %>%
  filter(DNA_result == "" | DNA_result == NA)

#2012 - BB, DNA test and eye test with age-adjusted phenotypic penetrance
At_Risk_2012 <- AF_Sub_TrendAge[AF_Sub_TrendAge$DOB > "2011-12-31",]
At_Risk_2012 <- At_Risk_2012[At_Risk_2012$DOB < "2013-01-01",]
At_Risk_2012 <- At_Risk_2012[!is.na(At_Risk_2012$IID),]
At_Risk_2012 <- At_Risk_2012 %>%
  filter(BB_geno > 0.0756)

x <- AF_Sub_Data$IRN %in% At_Risk_2012$IID 
y <- AF_Sub_Data[x,]
UnknownDNA <- y %>%
  filter(DNA_result == "" | DNA_result == NA)

#2013 - AB, DNA test and eye test with age-adjusted phenotypic penetrance
At_Risk_2012 <- AF_Sub_TrendAge[AF_Sub_TrendAge$DOB > "2011-12-31",]
At_Risk_2012 <- At_Risk_2012[At_Risk_2012$DOB < "2013-01-01",]
At_Risk_2012 <- At_Risk_2012[!is.na(At_Risk_2012$IID),]
At_Risk_2012 <- At_Risk_2012 %>%
  filter(AB_geno > 0.399)

x <- AF_Sub_Data$IRN %in% At_Risk_2012$IID 
y <- AF_Sub_Data[x,]
UnknownDNA <- y %>%
  filter(DNA_result == "" | DNA_result == NA)

```
